### 2022-04-20 response_reviewers.tex

\end{Reply}

(2) Adding on to the first point, how are the indicators measured? Should the conference organizers quantify the demographics of the attendees/participants?

\begin{Reply}

The evaluation of goals and outcomes is part of Rule 10, and has now been addressed in the manuscript in lines \ref{info} to \ref{info2}:
\textit{
(...) You will likely be able to evaluate the outcomes associated with these goals directly from the information collected from the organizing team, speakers, and participants (e.g.\ city/state/country of origin or preferred language).
You may have set goals concerning how welcomed and included participants felt during the conference, and need to assess if the implemented practices translated into true inclusion. 
Surveys, focus groups, and informal conversations during and after the conference are good ways to collect this information and great opportunities to receive feedback from the attendees.
}
We also added in Rule 1 an explicit link to Rule 10 in line \ref{rule10} to let the reader know that the evaluation results will be used as part of the conference wrap-up and continuity.

@@

As stated in Rule 10 (lines \ref{account-3} to \ref{data-share2}), there should be a final assessment of the achievement of the diversity and inclusion goals at the end of the conference. Many tools can be used for that (e.g.\ demographic information, surveys, focus groups, information from informal conversations). 
Publishing these final assessments is an important way to hold the organizers accountable both to the greater community and the committee itself.
\end{Reply}

(3) In Rule 4, it states `Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;' Does this pertain to invited speakers only, or speakers selected from submitted abstracts? If this pertains to scored abstracts, how does one ensure that this is done in a transparent and reproducible way? Or maybe I'm not fully grasping what curate means in this context?

\begin{Reply}

This specific section refers to the final selection of submitted proposals. Counteracting bias in invited roles was mentioned earlier in the text in lines \ref{invited} to \ref{invited-2}.
We agree with the reviewer that `curation' was an ambiguous term.
We rephrased the section as follows: \textit{
Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} (lines \ref{criteria-1}-\ref{criteria-2}).

@@

\end{Reply}

(4) In Rule 5, it states `The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.' How do the tablets help in-person attendees interact with remote participants? Especially participants that may be in a different time-zone. Or are the tablets used to help the online activities mentioned in the next paragraph?

\begin{Reply}

Although one can assume that an ever growing proportion of people have their own handheld device to interact virtually with remote participants, we cannot assume that everyone does, or that the models are compatible with the software platforms used in the conference. 
If the conference organizers provide such devices, Q\&A sessions and networking activities can be held virtually and allow for more integration between in-person and online participants. We edited the manuscript for clarity: \textit{The physical venue would have to include multiple ways to connect with online folks [...], and tablets to help in-person attendees without suitable personal devices to interact with remote participants during Q\&A or networking sessions} (line \ref{tablets}).

@@

(5) In Rule 7 it states `Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language.' It is not trivial to submit an abstract in two languages, and if there is no chance that the non-English abstract won't be reviewed I feel like this would be wasting their time. Is it possible to denote the accepted languages ahead of time? Or for the submitter to put in a language request before the abstract deadline?

\begin{Reply}

    Thank you for this suggestion. 
    The organizing team should decide whether to ask for two languages or just one, depending on their capabilities. 
    Indeed, it is not trivial to submit abstracts in two languages. However, from personal experience, we know that many people find English so cumbersome that, if given the choice, they would opt for adding a version in their native language to increase the chances of communicating their ideas accurately. 
    The text now states: 
    \textit{Allow for abstract submission in both English and the applicant's preferred native language, and whenever possible, assign a reviewer who is fluent in that language. 
    Alternatively, the organizing team could define a set of accepted languages for abstract submission, and allow applicants to choose from this list. 
    The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency} (lines \ref{abstractlanguage1} to \ref{abstractlanguage2}).

    We also added a mention to accepting multilingual submissions when counteracting self-selection during the call for abstracts in Rule 4, linking to Rule 7 (lines \ref{multilingual-submission} to \ref{multilingual-submission2}).
\end{Reply}

@@

    Thank you for the thorough and positive review. We hope to have addressed all of your concerns and particularly accountability. We agree with involving the community in every part of the process in an active and participative manner, using different channels of open communication and integrating members of the community into the organizing team. 
\end{Reply}

\section*{Reviewer \#2:}
L78-L80 As a female scientist I would ask myself "where are we" (because we still have parts of the World where female scientist are rare on leading positions), but I consider science apolitical in every segment, and I certainly suggest that it concentrate here on the scientific youth, the segment of research in industry, etc., and not on the above.
The rest of the paper it should certainly be praised because it enters into the essence of the exchange of knowledge today.

\begin{Reply}

    We thank the reviewer for this comment. The list given was meant to be an example of existing thematic groups and communities, aiming to go beyond gender representation.
    We have now added women as an example in line \ref{wis}:
    \textit{Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.}
\end{Reply}

@@

I have some comments for authors to consider, and possibly expand upon.

1. As much as I enjoyed reading the ten rules, I missed finding mention of how this is to be implemented and governed. Therefore, it is poignant that authors should brief discuss having an inclusion oversight or governance committee that tracks these metrics and their implementation.

\begin{Reply}

    A Diversity, Equity, and Inclusion (DEI) committee would be best suited to ensure that the inclusion vision and goals are achieved. 
    Establishing a DEI team is now suggested in Rule 3 in the manuscript (lines \ref{subteam-1} to \ref{subteam-2}):
    \textit{ 
    Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
    The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} 
    Thank you for pointing this out. 
    On the other hand, the existence of a DEI team is neither necessary nor sufficient, and we believe the most important thing is the commitment of the 
    whole organizing team towards inclusion, and the willingness to hold themselves accountable, with concrete actions mentioned in Rule 8 (during the conference, lines \ref{accountability-1} to \ref{accountability-2}) and Rule 10 (after the conference, lines \ref{account-3} to \ref{info2}) .

@@

    The coauthors are from diverse backgrounds, career stages, and different continents. They have experienced inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege, as stated in lines \ref{background-a}-\ref{background-b}.
    The manuscript summarizes our work together and is strongly supported by the awareness and experience of organizing a global conference that aims to be a kind, inclusive, and accessible experience for as many people as possible.
\end{Reply}

3. In Rule 1, though gender or ethnic diversity has been highlighted, geographical diversity needs to be discussed in detail. There is significant "north-south" divide in scientific research - with several of institutions as well as researchers from LMICs or global south not being given appropriate representation on conference committees as well as on speakers diversity.

\begin{Reply}

   On a global scale, there is an imbalance associated with what could be considered a `north-south' divide, though the definition of north and south may not be clear. We agree that coming from a lower income country hinders participation in conferences. For that reason, we offer an example of registration rates based on country of origin and career status (https://spcanelon.github.io/useR2021-cost-conversion-tool/). 
   On the other hand, while the manuscript is heavily based on our experience organizing a global conference—which is why there are some global examples—the intention is to make the rules general and flexible enough to be useful for local or regional conferences around the world. 
   For that reason, we explicitly mention in Rule 1 that \textit{Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference} (line \ref{indicators-1}).

### 2022-04-20 clean.tex
Response submitted :sunglasses:

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
\hl{* Translated versions of the extended abstract are available in 10 languages in the Supplementary Information--Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} \hl{The rules were first drafted by the core team and members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

\linelabel{indicators-1}\hl{
Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
Assess how it translates to the conference's particular field and scientific or professional community.
Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if they exist. 
Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.} 
\linelabel{indicators-2}
Define measurable outcomes for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions). 
They will help you be accountable along the way (see \textbf{Rule 10}). \linelabel{rule10} %\hl{and assess to what extent your inclusion goals were met (see \textbf{Rule 10})}.
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

@@

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part in the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects included in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it is good to have them in mind from the beginning.
Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM\hl{, women in science}\linelabel{wis}), and sessions aimed to welcome first-timers to the conference.

% A culture of support (active bystandership) and code of conduct for a safe place 
Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, inappropriate, or offensive happens.
This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
However, the most important step that you should take to ensure the safety of the community is
adopting a code of conduct (CoC) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
\linelabel{unacceptable-1}The CoC is a living document meant to keep the community safe and should state clearly the behaviors that are deemed unacceptable by the community, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}. 
\hl{Bear in mind that every dimension of diversity can be a target of unacceptable behavior. The CoC should help honor the vision expressed in the diversity statement.}\linelabel{unacceptable-2}

The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media, or in organization meetings).  
Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding funds for their work to your budget (see \textbf{Rule 9}).
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.

@@

% Build a diverse team, and representation
A genuinely inclusive conference can only be organized by an inclusive and diverse team.
Organizers should comprise a team with people from as many different backgrounds as possible in terms of regions, genders, ethnicities, career stages, and other aspects of diversity.
People with disabilities often say: `Nothing about us without us' \cite{charlton_nothing_1998, werner_nothing_1998}; the same holds for other dimensions of diversity. 
Having people from diverse backgrounds in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
The experience of people in marginalized groups
is especially important; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupying critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
\hl{If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into.}

\linelabel{subteam-1} \hl{Create sub-teams to distribute fairly the large amount of work that organizing a conference represents. 
Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
The core team should support the DEI initiatives and advocate for inclusion across all conference areas.} \linelabel{subteam-2} 
It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
every member should have the freedom to choose which areas of the conference they want to contribute to.
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
In any case, try to find meaningful ways to acknowledge everybody's work. Use the conference spaces (e.g. webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other for the long haul.

@@

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate such self-selection, some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
\linelabel{multilingual-submission}\hl{To mitigate self-selection from non-native speakers, consider accepting submissions in several languages} (see \textbf{Rule 7}).

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias} 
Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing \hl{the submission information as much as possible} \linelabel{anonymous}, and
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
\linelabel{criteria-1} \hl{Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.} \linelabel{criteria-2}
\linelabel{video-1}\hl{It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others.} \cite{spitschanVideoGrantProposals2021}\linelabel{video-2}

@@

If you are organizing a hybrid event, decide on the relative importance that online and in-person components will take \cite{bajpai_towards_2021}. The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending the conference.
The audience, chairs, and presenters should all be able to interact regardless of their in-person or remote status. 
To best integrate in-person and remote attendees, all questions and comments should pass through a single online system. 
The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants.

@@

% Rocío: main idea: make conferences more linguistically inclusive
Create an inclusive environment by encouraging the full participation of non-native English speakers. This may be done in multiple ways.
\linelabel{abstractlanguage1}\hl{Allow for abstract submission in both English and the applicant's preferred native language, and whenever possible, assign a reviewer who is fluent in that language. 
Alternatively, the organizing team could define a set of accepted languages for abstract submission, and allow applicants to choose from this list. 
The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency}.\linelabel{abstractlanguage2}
Whenever possible, identify other prominent languages for the conference and provide translated captions or live language interpretation into these key languages, including multilingual Q\&A sessions.
Another way to engage non-native English participants is to host sessions and events in languages other than English. 
Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same importance as the rest of the program.
Importantly, consideration and respect of accents or linguistic mistakes can make a significant difference in the conference experience of non-native English speakers. Kindly remind your native English-speaking audience to be mindful of that.

@@

The communication strategy of your conference should aim to reach broader audiences and express the commitment to diversity and inclusion and the welcoming spirit of the conference.
Actively reach out and promote the conference to communities that have been systematically excluded. 
Promote the abstract submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).
\linelabel{promotion-1}\hl{During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility as all other sessions.}\linelabel{promotion-2}

\linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
\hl{These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.}
Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues. 
For instance, inform participants if the conference platform is not completely screen-reader friendly and let people know if you are offering help to navigate it, or highlight on the conference website if captions will be available for some talks but not all.
Provide a point of contact to help clarify any questions regarding accessibility, financial support, and other issues affecting participation.
\linelabel{accountability-2}

>>>>

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

\linelabel{indicators-1}
Lack of diversity will vary across areas of knowledge and depend on the geographic context of the conference.
Assess how it translates to the conference's particular field and scientific or professional community.
Some information can come from Diversity, Equity, and Inclusion (DEI) studies in the field or at a larger scale (e.g. STEM---science, technology, engineering, and mathematics) if they exist. 
Most importantly, involve the community---the subset of the general public that would ideally attend the conference---to help make this assessment and set the vision and goals of diversity and inclusion for the conference.
\linelabel{indicators-2}
Define measurable outcomes for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions). 
They will help you be accountable along the way (see \textbf{Rule 10}).
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

@@

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part in the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects included in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it is good to have them in mind from the beginning.
Open the schedule for events organized by thematic groups and communities (e.g.\ task force meetings, LGBTQIA+-friendly spaces, Black in STEM, women in science), and sessions aimed to welcome first-timers to the conference.

% A culture of support (active bystandership) and code of conduct for a safe place 
Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, inappropriate, or offensive happens.
This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
However, the most important step that you should take to ensure the safety of the community is
adopting a code of conduct (CoC) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
\linelabel{unacceptable-1}The CoC is a living document meant to keep the community safe and should state clearly the behaviors that are deemed unacceptable by the community, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}. 
Bear in mind that every dimension of diversity can be a target of unacceptable behavior. The CoC should help honor the vision expressed in the diversity statement.

The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because CoC violations can happen before the conference begins (e.g.\ on social media or in organization meetings).  
Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding funds for their work to your budget (see \textbf{Rule 9}).
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.

@@

% Build a diverse team, and representation
A genuinely inclusive conference can only be organized by an inclusive and diverse team.
Organizers should comprise a team with people from as many different backgrounds as possible in terms of regions, genders, ethnicities, career stages, and other aspects of diversity.
People with disabilities often say: `Nothing about us without us' \cite{charlton_nothing_1998, werner_nothing_1998}; the same holds for other dimensions of diversity. 
Having people from diverse backgrounds in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
The experience of people in marginalized groups
is especially important; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupying critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into.

Create sub-teams to distribute fairly the large amount of work that organizing a conference represents. 
Having a dedicated DEI team may be an excellent way to give the topic the attention and energy it requires.
The core team should support the DEI initiatives and advocate for inclusion across all conference areas.
It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
every person should have the freedom to choose which areas of the conference they want to contribute to.
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
Instead, nominate folks directly and offer them leadership positions in the organizing team, especially those positions that would be typically occupied by people from privileged groups.
Moreover, some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
In any case, try to find meaningful ways to acknowledge everybody's work. Use the conference spaces (e.g. webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other for the long haul.

@@

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate such self-selection, some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.
To mitigate self-selection from non-native speakers, consider accepting submissions in several languages (see \textbf{Rule 7}).

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted.\linelabel{bias} 
Selection committees need to be reminded of such biases in the evaluation process, to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing the submission information as much as possible and
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
Even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society. It is, therefore, essential to examine the final list of selected abstracts to check if there is a lack of diversity. If proposals with similar quality were still judged differently because of unconscious bias, consider giving preference to the work from people in minoritized groups.
This manual examination at the final stage is unlikely to be reproducible. Nonetheless, the whole process should be as transparent as possible: evaluation criteria should be clearly stated and shared with the reviewers and authors, preferably during the call for abstracts.
It is worth noting that self-selection and unconscious bias may be aggravated by the use of video abstracts as an alternative to written ones, affecting people without the resources to create good quality videos, people with disabilities, non-native speakers, racialized people, and people with diverse body types, among others. \cite{spitschanVideoGrantProposals2021}

@@

If you are organizing a hybrid event, decide on the relative importance that online and in-person components will take \cite{bajpai_towards_2021}. The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending the conference.
The audience, chairs, and presenters should all be able to interact regardless of their in-person or remote status. 
To best integrate in-person and remote attendees, all questions and comments should pass through a single online system. 
The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees without suitable personal devices to interact with remote participants during Q\&A or networking sessions.

@@

% Rocío: main idea: make conferences more linguistically inclusive
Create an inclusive environment by encouraging the full participation of non-native English speakers. This may be done in multiple ways.
Allow for abstract submission in both English and the applicant's preferred native language, and whenever possible, assign a reviewer who is fluent in that language. 
Alternatively, the organizing team could define a set of accepted languages for abstract submission, and allow applicants to choose from this list. 
The goal should be to judge the abstracts primarily by the quality or relevance of the work instead of English proficiency.
Whenever possible, identify other prominent languages for the conference and provide translated captions or live language interpretation into these key languages, including multilingual Q\&A sessions.
Another way to engage non-native English participants is to host sessions and events in languages other than English. 
Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same importance as the rest of the program.
Importantly, consideration and respect of accents or linguistic mistakes can make a significant difference in the conference experience of non-native English speakers. Kindly remind your native English-speaking audience to be mindful of that.

@@

The communication strategy of your conference should aim to reach broader audiences and express the commitment to diversity and inclusion and the welcoming spirit of the conference.
Actively reach out and promote the conference to communities that have been systematically excluded. 
Promote the abstract submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).
During the conference, pay special attention to promoting the sessions led, chaired, and presented by people in minoritized groups, multilingual sessions, and the diversified thematic sessions (see \textbf{Rule 4}) to give them the same visibility as all other sessions.

\linelabel{accountability-1}Publish the diversity statement, code of conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9}).
These guidelines will not only communicate the conference values and practices but will also act as a built-in mechanism to hold the organizers accountable.
Be transparent with potential attendees, communicate the limitations of the conference to let them know what to expect, and outline ways in which the organizing team will try to mitigate these issues. 
For instance, inform participants if the conference platform is not completely screen-reader friendly and let people know if you are offering help to navigate it, or highlight on the conference website if captions will be available for some talks but not all.
Provide a point of contact to help clarify any questions regarding accessibility, financial support, and other issues affecting participation.
\linelabel{accountability-2}
